---
title: "Naive Bayes Lab"
author: "Ilana Radinsky"
date: "July 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises

1. Suppose you have a database on four customers. You know their income and whether or not they bought your product. Create a data frame with this data.

```{r}
train <- data.frame(buy=c("yes","no","no","yes"), 
                    income=c("high","high","medium","low"))

train
```

2. Using Bayes rule calculate the probability that a customer will buy your product given that he or she has high income.

P(buy==yes|high) = [P(high|buy==yes)*P(buy==yes)]/P(high)
P(buy==yes|high) = [(1/2)*(2/4)] / (2/4) = 1/2


3. Estimate naive Bayes model using your data above. What is the prior probability of someone buying your product? What is the probability that a customer has a high income give that he or she bought your product?

```{r pressure, echo=FALSE}
library(e1071)

classifier <- naiveBayes(buy ~ income, train)
classifier
```

4. Using the model you estimated above predict the probability of buying given that a customer has high income. Is your result the same as the one you calculated by hand in question 1?

```{r}
test <- data.frame(buy=c("yes"), income=c("high"))
test$buy <- factor(test$buy, levels=c("yes","no"))
test$income <- factor(test$income, levels=c("high", "low", "medium"))
test

prediction <- predict(classifier, test, type="raw")
prediction
```

5. Suppose you have a database on four customers. You know their gender, income and whether or not they bought your product. Create a data frame with this data.

```{r}
train2 <- data.frame(buy=c("yes","no","no","yes"), 
                    income=c("high","high","medium","low"),
                    gender=c("male", "female", "female", "male"))
train2
```

6. Using Bayes rule calculate the probability that a customer will buy your product given that he has high income and male.

P(buy==yes|high, male) = [P(high, male|buy==yes)*P(buy==yes)]/P(high, male)
P(buy==yes|high, male) = [(1/2)*(1/2)] / (1/4) = 1


7. Estimate naive Bayes model using your data above. What is the prior probability of someone buying your product? What is the probability that a customer has a high income given that he or she bought your product? What is the probability that a customer is male given that he bought your product?

```{r}
classifier2 <- naiveBayes(buy ~ income + gender, train2)
classifier2
```

8. Using the model you estimated above, predict the probability of buying given that a customer has a high income and is male. Is your result the same as the one you calculated by hand in question 1?

```{r}
test2 <- data.frame(buy=c("yes"), income=c("high"), gender=c("male"))
test2$buy <- factor(test2$buy, levels=c("yes","no"))
test2$income <- factor(test2$income, levels=c("high", "low", "medium"))
test2$gender <- factor(test2$gender, levels=c("male", "female"))
test2

prediction2 <- predict(classifier2, test2, type="raw")
prediction2

```

































































