---
title: "Writeup"
author: "Ilana Radinsky"
date: "June 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
load("outputs.RData")
```

## Introduction
This is a template for a write-up. You can see more advanced ones [here](http://rmarkdown.rstudio.com/developer_document_templates.html) and [here](http://rmarkdown.rstudio.com/gallery.html).

## Data
We used data from ....

## Anlyses

1. Which genre is the highest rated?


```{r, echo=FALSE, messages=FALSE}
plot1
```

2. Which genre is the most popular?

```{r, echo=FALSE, messages=FALSE}
plot2
```

3. How does appr/enjoyment with movie gross income?

```{r, echo=FALSE, messages=FALSE}
plot3
```

4. Are older movies more likely to be considered meaningful?

```{r, echo=FALSE, messages=FALSE}
plot4
```

5. Are "classic" movies more likely to be considered meaningful?

```{r, echo=FALSE, messages=FALSE}
plot5
```

## Discussion
We think we did a great job.

## Appendix
More details about the data.

We registered the study at ....

